Robust Adaptation to Multi-Scale Climate Variability

### 2019-05-17 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
Update paper-2018-robust-adaptation-cyclical-risk.tex
Final version

% Some package settings
\sisetup{round-mode=figures,round-precision=3,scientific-notation=false}
\graphicspath{{./figs/}{../fig/}} % folders of figures

@@

% Key Points
\begin{keypoints}
  \item \add{Quasi-periodic and secular climate signals, with different identifiability and predictability, control future uncertainty and risk}
  \item \add{Adaptation strategies need to consider how uncertainties in risk projections influence success of decision pathways}
  \item \add{Stylized experiments reveal how bias and variance of climate risk projections influence risk mitigation over a finite planning period}
\end{keypoints}

% Abstract
\begin{abstract}
  The assessment and implementation of structural or financial instruments for climate risk mitigation requires projections of future climate risk over the operational life of each proposed instrument.
  A point often neglected in the climate adaptation literature is that the physical sources of predictability differ between projects with long and short planning periods: while historical and paleo climate records \add{emphasize} \remove{highlight the role of low-frequency variability (LFV) in modulating climate extremes at} interannual to multidecadal \remove{time scales} \add{modes of variability}, \acrlong{acc} is expected to alter their occurrence at longer time scales.
  In this paper we present a set of stylized experiments to assess the uncertainties and biases involved in estimating future climate risk over a finite future period, given a limited observational record.
  These experiments consider both quasi-periodic and secular change for the underlying risk, as well as statistical models for estimating this risk from an $N$-year historical record.
  The uncertainty of \acrshort{ipcc}-like future scenarios is considered through an equivalent sample size $N$. 
  The relative importance of estimating the short- or long-term risk extremes depends on the investment life $M$\remove{ and the future discount rate for investment}.
  Shorter design lives are preferred \add{for situations} where inter-annual to decadal variability can be successfully identified and predicted, suggesting the importance of sequential investment strategies for adaptation. 
\end{abstract}

@@

Recent climate extremes such as floods, droughts, hurricanes, tornadoes, hailstorms, and heat waves have caused death and destruction, motivating investments in climate adaptation for the public and private sectors.
Further, rapid and continuing changes to global climate hazard and exposure underscore the need for adaptation strategies.
For example, population growth and urbanization have driven rapid increases in global exposure to events such as floods \citep{Jongman:2012cr} and tropical cyclones \citep{Peduzzi:2012iq}.
At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content necessitate the development of} novel adaptation strategies\remove{ are needed}.

This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, e.g.\@\xspace,][]{Merz:2014gf}.
\add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
Although climate risk has traditionally been managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently \remove{slowed their implementation}\add{dampened enthusiasm}.
\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, a catastrophe bond)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
These instruments are not typically implemented in isolation or statically.
Instead, investment decisions made at each point in time affect the viability, costs, and benefits of future decisions, causing the system to trace a ``pathway'' through time \citep{Walker:2013gi,Haasnoot:2013im,Haasnoot:2015dh}.

Despite recent insights, important questions remain.
How should a portfolio of risk mitigation instruments be optimized?
How should one choose between permanent and transient instruments?
Under what conditions is a permanent, large infrastructure investment required, and what information is needed to recognize this threshold?
In this paper we focus on \add{the narrower question of} how the temporal structure of climate risk, and the uncertainties associated with \add{its estimation}\remove{estimating it}, influence the answers to these questions.
We continue this section with three specific observations about climate risk which, while \remove{apparently}\add{seemingly} obvious, have important and subtle implications \add{that}\remove{which} we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.

@@

Typical climate risk management policies do not use a single risk mitigation instrument, but rather build a portfolio of several instruments.
Each has its own operational period, which may or may not match the planning horizon of the portfolio as a whole.
This means that even if the portfolio has a long planning period, i.e.\@\xspace if long-term plans are a priority, this goal may be best accomplished through a series of flexible and adaptive instruments with short individual planning periods.
\add{For example, the optimal policy for New York City to manage uncertain hurricane risk in the 21st century might potentially be to keep areas devastated by hurricane Sandy zoned for low-impact development for the next 10 years.
This would reduce future risk over all climate scenarios while postponing major investments until large uncertainties as to the magnitude of future sea level rise are resolved.}
The costs and benefits of each individual instrument will be assessed over its individual, finite planning period, but decisions about the portfolio structure are evaluated over the longer planning horizon.

@@

Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
The Hurst phenomenon is well known and identified with long memory processes in geophysics, physics, biology, medicine, traffic, network dynamics, and finance \citep{oconnell:2016}.
The extensive observations of such behavior in hydrologic and climatic time series emphasize the need to consider such processes as underlying any discussion of climate change or nonstationarity \citep{koutsoyiannis:2003,markonis:2013,Palmer:1993ef}.
The Hurst phenomenon has also been connected to low frequency quasi-periodic phenomenon, especially where fractal scaling is expected.
For example, wavelet methods have been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.
The Hurst phenomenon also provides a link between catchment hydrology and global climate dynamics \citep{bloschl:2010,montanari:2003}.
The Hurst exponent is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
These processes have also been used to describe multi-scale dynamics of the climate \citep{lovejoy:2012,lovejoy:2013a,lovejoy:2013,selvam:2017}, including \gls{enso} \citep{maruyama:2018,zivkovic:2013} and the \gls{pdo} \citep{Mantua:1997kj}.

External forcing from structured climate signals \citep[``teleconnections'';][]{Angstrom:1935ej} and catchment dynamics are both useful in explaining the \gls{lfv} observed in natural hydroclimate time series.
We illustrate such \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of annual maximum streamflowflow on the American River at Folsom, and the global wavelet power spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
Peaks for the American River time series are apparent at 2.3 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
This is illustrated by the blue line in \remove{the top left panel of} \cref{fig:observed-lfv}\add{(b)}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
A detailed analysis of these time series is beyond the scope of this paper, but we note that the high amplitude and long time periods of the quasi-periodic oscillations they exhibit are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Kiem:2002kq,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw}.
The key implication is that the observations, \citep{Jain:2001hz}, trends \citep{bhattacharya:1983}, and frequencies \citep{newman:2016} observed in the past are often poor predictors of future behavior.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{observed-lfv.pdf}
  \caption{
    Hydroclimate time series vary on many time scales.
    (a) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda}.
    \remove{High numbers}\add{Lower values} indicate more severe drought.
    A 20-year running mean is also shown in blue.
    (b) A \SI{100}{year} record of annual-maximum streamflow for the American River at Folsom.
    Daily streamflow values were divided by the catchment area to yield a normalized flow in units of \si{\milli\meter\per\day}.
    (c) The global wavelet power spectrum of the \gls{lbda} time series (a).
    Blue (red) dots indicate frequencies which are significant at $\alpha=0.10(0.05)$ compared to white noise.
    (d) Global wavelet power spectrum, like (c), for the American River data.
  }\label{fig:observed-lfv}
\end{figure}

@@

Evaluating a particular risk mitigation instrument involves projecting climate risk over the $M$-year planning period.
Consequently, the physical mechanisms which impart predictability on the system differ between projects with long and short planning periods.
As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a \SI{100}{year} planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
Even \remove{with} \add{in the idealized and unrealistic case of} a perfect climate model, these uncertainties will be large.
By contrast, this perfect climate model may usefully inform estimates of climate hazard over a three-year insurance contract with much less associated uncertainty.

Of course, scientists are not equipped with perfect models.
Since different physical processes control climate risk at different timescales, successful integration of climate projections into decision frameworks depends on identifying, and subsequently predicting, these processes.
A key question is whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portend\remove{s} for risk mitigation instruments with a planning period ranging from a few years to several decades.
As shown in \cref{fig:conceptual-sketch} (b), the combination of \gls{lfv}, stochastic variability, and secular change in a limited record can lead to large uncertainty in estimated future risk.
Although \cref{fig:conceptual-sketch} focuses on physical processes, similar conclusions would also be valid for the socioeconomic processes which drive exposure to floods and other hydroclimate hazards.

@@

We consider three scenarios for climate risk, which we define by the structure of the underlying climate \annote{signal}{changed the order}:
\begin{enumerate*}[label= (\roman*)]
  \item secular change only;
  \item \gls{lfv} only; and
  \item \gls{lfv} plus secular change.
\end{enumerate*}
For each scenario, and for its identification from the $N$ year length historical data, the bias and variance of the estimated flood risk over the $M$ year design life relative to the ``true model'' are computed.
We repeat the simulations $J = 1000$ times for each combination of experiment parameters to obtain estimates of the expected bias and variance for each scenario given $M$ and $N$ (\cref{sec:methods-evaluating}).

@@

The first step is to sample climate risk by generating synthetic streamflow sequences.
To do this, we model  annual-maximum flood peaks with a log-normal distribution, conditional on a location parameter which varies in time:
\begin{equation} \label{eq:log-normal}
  \log Q(t) \sim \mathcal{N} \qty(\mu(t), \sigma(t)).
\end{equation}
We further assume a constant coefficient of variation of the log streamflow,
\begin{equation}
  \sigma(t) = \xi \mu(t)  
\end{equation}
and apply a lower threshold on the standard deviation
\begin{equation}
  \sigma(t) \geq \sigma_\text{min} > 0.
\end{equation}
This formulation describes \add{all} scenarios \add{for future climate considered in this paper within} \remove{in} a single equation\remove{;}\add{.}
\add{To add climate variability to the system,} the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
\begin{equation}\label{eq:nino3}
  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
\end{equation}
where $x(t)$ represents a climate time series which itself exhibits \gls{lfv} but not secular change.
This parameterization is analagous to the ``climate-informed'' approach described in several studies for estimating climate risk \citep{Delgado:2014ey,Merz:2014gf,Farnham:2018gs}.
Following \cref{eq:nino3}, when $\beta \neq 0$ there will be \gls{lfv}, and when $\gamma \neq 0$ there will be secular change.
The values of all parameters used for sampling climate risk are listed in the online supporting information for each of the three scenarios considered.

We represent the climate state variable $x(t)$ through an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} as well as a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
We model \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
To create an annual time series, we average the October-December values of the NINO3 index for each year.
Supplemental figure S1 shows a wavelet spectrum and time series plot of the resulting annual time series.
In the online supplemental information we consider an alternative parameterization of $\mu(t)$, \remove{i.e.\@\xspace an alternative to,} which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.

@@

Once a synthetic streamflow sequence has been generated, we evaluate the identifiability and predictability of the dominant climate modes by fitting the sequence to statistical models and creating probabilistic projections of the future.
We use three well-studied statistical methods for future flood risk, each of which parameterizes time in a different way.
One is purely stationary, another captures \gls{lfv}, and the third captures secular change.
We choose these models for their interpretability and simplicity, rather than because of a belief that they are generally valid.
For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations.
Once a statistical model is fit to these \remove{``}observations\remove{''}, then $K=\num{1000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.

\add{In the first case we} \remove{We first consider fitting}\add{fit} a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
\add{We refer to this model as ``LN2 Stationary.''}
The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors \citep{gelman:2017,simpson:2017}.
The full model, including priors, is given in supplemental equation (S1).

Next, we modify this stationary model to incorporate secular change.
Many studies have done this by regressing certain parameters of the model on time \citep[see][for a comprehensive review]{Salas:2018ge}.
We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in supplemental equation (S2).
\add{We refer to this model as ``LN2 Linear Trend''}
This model gives a lower bound on total informational uncertainty because it correctly represents the trend's known form, whereas in real-world analyses the form of the trend is unknown.

\remove{We finally consider}\add{Finally, we explicitly model} \remove{a HMM which explicitly represents}\gls{lfv} \add{using a} \gls{hmm}.
An \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (i.e.\@\xspace hidden) states $S(t)$ \citep{Rabiner:1986jk}.
The (unobserved) states evolve following a first-order Markov process, and the observed variable (e.g.\@\xspace streamflow) depends only on the underlying state.
\Glspl{hmm} have been widely used for modeling streamflow sequences \citep{Bracken:2016ba} and \gls{enso} \citep{rojohernandez:2017}.
We fit  streamflow sequences using a \gls{hmm} with two states.
The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution \remove{conditional only on the state}\add{that is conditional only on the unobserved state variables}.
This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
For simplicity, we fit only a two-state \gls{hmm} to each sequence.
Future floods are then estimated by simulating future states from the estimated transition matrix and then drawing $Q(t)$ conditional on the simulated state.

@@

Both estimation bias and estimation uncertainty affect the utility of a climate risk projection.
An instrument \add{whose design was based on projections with} \remove{designed based on} overestimated variance \remove{and}\add{or} positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the \remove{user}\add{public}.

We evaluate both the estimation bias and estimation uncertainty.
For a given choice of $M$, $N$, \add{and} generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'' and the $K=1000$ posterior simulations of future flows.
The quantity $\hat{p}_T$, the estimated expected number of floods per year, is taken by calculating, for each of the $K$ posterior simulations, the number of exceedances of the flood design threshold, then dividing by $M$ to get exceedances per year.
We then compute the variance of these $K$ estimates.
We further calculate the bias of $\hat{p}_T$ by averaging it across the $K$ samples and comparing this to the number of times the $M$-year ``future period'' of the synthetic streamflow sequence exceeds the flood design threshold.
Since the ``observed'' number of flood exceedances from the generating model is inherently noisy for an $M$-year period, we average the bias and variance across $J=1000$ different streamflow sequences to compute expected values of both.
These sequences are generated with the same underlying parameters, but the specific synthetic NINO3 sequence (or set of Markov states) may differ between the $J$ sequences.

@@

\Cref{fig:flowchart} describes the experimental design.
We assess estimation bias and variance for three \remove{cases}\add{scenarios of future climate}.
First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped (``secular change only'').
Next, we consider the ``pre-industrial'' case where there is no secular change but \gls{lfv} modulates climate risk in time (``\acrlong{lfv} only'').
Finally, we consider a more realistic (though still idealized) case with both \gls{lfv} and secular change (``\acrlong{lfv} plus secular change'').
Model parameters for each scenario are given in the supplemental methods section of the online supporting materials.

>>>>

% Some package settings
\sisetup{round-mode=figures,round-precision=3,scientific-notation=false}
%\graphicspath{{./}} % folders of figures

\graphicspath{{./figs/}{../fig/}} % folders of figures

@@

% Key Points
\begin{keypoints}
  \item Quasi-periodic and secular climate signals, with different identifiability and predictability, control future uncertainty and risk
  \item Adaptation strategies need to consider how uncertainties in risk projections influence success of decision pathways
  \item Stylized experiments reveal how bias and variance of climate risk projections influence risk mitigation over a finite planning period
\end{keypoints}

% Abstract
\begin{abstract}
  The assessment and implementation of structural or financial instruments for climate risk mitigation requires projections of future climate risk over the operational life of each proposed instrument.
  A point often neglected in the climate adaptation literature is that the physical sources of predictability differ between projects with long and short planning periods: while historical and paleo climate records emphasizemodes of variability, \acrlong{acc} is expected to alter their occurrence at longer time scales.
  In this paper we present a set of stylized experiments to assess the uncertainties and biases involved in estimating future climate risk over a finite future period, given a limited observational record.
  These experiments consider both quasi-periodic and secular change for the underlying risk, as well as statistical models for estimating this risk from an $N$-year historical record.
  The uncertainty of \acrshort{ipcc}-like future scenarios is considered through an equivalent sample size $N$. 
  The relative importance of estimating the short- or long-term risk extremes depends on the investment life $M$.
  Shorter design lives are preferred for situations where inter-annual to decadal variability can be successfully identified and predicted, suggesting the importance of sequential investment strategies for adaptation.
\end{abstract}

@@

Recent climate extremes such as floods, droughts, hurricanes, tornadoes, hailstorms, and heat waves have caused death and destruction, motivating investments in climate adaptation for the public and private sectors.
Further, rapid and continuing changes to global climate hazard and exposure underscore the need for adaptation strategies.
For example, population growth and urbanization have driven rapid increases in global exposure to events such as floods \citep{Jongman:2012cr} and tropical cyclones \citep{Peduzzi:2012iq}.
At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
Even if future mitigation efforts are successful, existing levels of atmospheric $\text{CO}_2$ and ocean heat content necessitate the development of novel adaptation strategies.

This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, e.g.\@\xspace,][]{Merz:2014gf}.
A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures \citep{lempert:2007,Borgomeo:2018hk} and which fail along non-catastrophic modes \citep{brown:2010}.
Although climate risk has traditionally been managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently dampened enthusiasm.
Rather, actors such as New York City have turned to a combination of structural (e.g.\@\xspace, stormwater barrier), operational (e.g.\@\xspace, improved evacuation routes), and financial (e.g.\@\xspace, a catastrophe bond) instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
These instruments are not typically implemented in isolation or statically.
Instead, investment decisions made at each point in time affect the viability, costs, and benefits of future decisions, causing the system to trace a ``pathway'' through time \citep{Walker:2013gi,Haasnoot:2013im,Haasnoot:2015dh}.

Despite recent insights, important questions remain.
How should a portfolio of risk mitigation instruments be optimized?
How should one choose between permanent and transient instruments?
Under what conditions is a permanent, large infrastructure investment required, and what information is needed to recognize this threshold?
In this paper we focus \add{more narrowly on the temporal}\remove{ on the narrower question of how the temporal} structure of climate risk, and the uncertainties associated with its estimation, influence the answers to these questions.
We continue this section with three specific observations about climate risk which, while seemingly obvious, have important and subtle implications that we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.

@@

Typical climate risk management policies do not use a single risk mitigation instrument, but rather build a portfolio of several instruments.
Each has its own operational period, which may or may not match the planning horizon of the portfolio as a whole.
This means that even if the portfolio has a long planning period, i.e.\@\xspace if long-term plans are a priority, this goal may be best accomplished through a series of flexible and adaptive instruments with short individual planning periods.
For example, the optimal policy for New York City to manage uncertain hurricane risk in the 21st century might potentially be to keep areas devastated by hurricane Sandy zoned for low-impact development for the next 10 years.
This would reduce future risk over all climate scenarios while postponing major investments until large uncertainties as to the magnitude of future sea level rise are resolved.
The costs and benefits of each individual instrument will be assessed over its individual, finite planning period, but decisions about the portfolio structure are evaluated over the longer planning horizon.

@@

Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
The Hurst phenomenon is \add{a} well known \add{mathematical relationship which describes the} \remove{and identified with} long memory \add{of} processes \add{found in} in geophysics, physics, biology, medicine, traffic, network dynamics, and finance \citep{oconnell:2016}.
The extensive observations of such behavior in hydrologic and climatic time series emphasize the need to consider such processes as underlying any discussion of climate change or nonstationarity \citep{koutsoyiannis:2003,markonis:2013,Palmer:1993ef}.
The Hurst phenomenon has also been connected to low frequency quasi-periodic phenomenon, especially where fractal scaling is expected.
For example, wavelet methods have been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.
The Hurst phenomenon also provides a link between catchment hydrology and global climate dynamics \citep{bloschl:2010,montanari:2003}.
The Hurst exponent is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
These processes have also been used to describe multi-scale dynamics of the climate \citep{lovejoy:2012,lovejoy:2013a,lovejoy:2013,selvam:2017}, including \gls{enso} \citep{maruyama:2018,zivkovic:2013} and the \gls{pdo} \citep{Mantua:1997kj}.

External forcing from structured climate signals \citep[``teleconnections'';][]{Angstrom:1935ej} and catchment dynamics are both useful in explaining the \gls{lfv} observed in natural hydroclimate time series.
We illustrate such \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of annual maximum streamflowflow on the American River at Folsom, and the global wavelet power spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
Peaks for the American River time series are apparent at 2.3 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
This is illustrated by the blue line in \cref{fig:observed-lfv}(b), which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
A detailed analysis of these time series is beyond the scope of this paper, but we note that the high amplitude and long time periods of the quasi-periodic oscillations they exhibit are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Kiem:2002kq,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw}.
The key implication is that the observations, \citep{Jain:2001hz}, trends \citep{bhattacharya:1983}, and frequencies \citep{newman:2016} observed in the past are often poor predictors of future behavior.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{observed-lfv.pdf}
  \caption{
    Hydroclimate time series vary on many time scales.
    (a) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda}.
    Lower values indicate more severe drought.
    A 20-year running mean is also shown in blue.
    (b) A \SI{100}{year} record of annual-maximum streamflow for the American River at Folsom.
    Daily streamflow values were divided by the catchment area to yield a normalized flow in units of \si{\milli\meter\per\day}.
    (c) The global wavelet power spectrum of the \gls{lbda} time series (a).
    Blue (red) dots indicate frequencies which are significant at $\alpha=0.10(0.05)$ compared to white noise.
    (d) Global wavelet power spectrum, like (c), for the American River data.
  }\label{fig:observed-lfv}
\end{figure}

@@

Evaluating a particular risk mitigation instrument involves projecting climate risk over the $M$-year planning period.
Consequently, the physical mechanisms which impart predictability on the system differ between projects with long and short planning periods.
As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a \SI{100}{year} planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
Even in the idealized and unrealistic case of a perfect climate model, these uncertainties will be large.
By contrast, this perfect climate model may usefully inform estimates of climate hazard over a three-year insurance contract with much less associated uncertainty.

Of course, scientists are not equipped with perfect models.
Since different physical processes control climate risk at different timescales, successful integration of climate projections into decision frameworks depends on identifying, and subsequently predicting, these processes.
A key question is whether the limited information in an $N$-year observational record permits the identification and projection of cyclical climate variability and secular change, and what the resulting bias and uncertainty portend for risk mitigation instruments with a planning period ranging from a few years to several decades.
As shown in \cref{fig:conceptual-sketch} (b), the combination of \gls{lfv}, stochastic variability, and secular change in a limited record can lead to large uncertainty in estimated future risk.
Although \cref{fig:conceptual-sketch} focuses on physical processes, similar conclusions would also be valid for the socioeconomic processes which drive exposure to floods and other hydroclimate hazards.

@@

We consider three scenarios for climate risk, which we define by the structure of the underlying climate signal:
\begin{enumerate*}[label= (\roman*)]
  \item secular change only;
  \item \gls{lfv} only; and
  \item \gls{lfv} plus secular change.
\end{enumerate*}
For each scenario, and for its identification from the $N$ year length historical data, the bias and variance of the estimated flood risk over the $M$ year design life relative to the ``true model'' are computed.
We repeat the simulations $J = 1000$ times for each combination of experiment parameters to obtain estimates of the expected bias and variance for each scenario given $M$ and $N$ (\cref{sec:methods-evaluating}).

@@

The first step is to sample climate risk by generating synthetic streamflow sequences.
To do this, we model  annual-maximum flood peaks with a log-normal distribution, conditional on a location parameter which varies in time:
\begin{equation} \label{eq:log-normal}
  \log Q(t) \sim \mathcal{N} \qty(\mu(t), \sigma(t)).
\end{equation}
We further assume a constant coefficient of variation of the log streamflow,
\begin{equation}
  \sigma(t) = \xi \mu(t)  
\end{equation}
and apply a lower threshold on the standard deviation
\begin{equation}
  \sigma(t) \geq \sigma_\text{min} > 0.
\end{equation}
This formulation describes all scenarios for future climate considered in this paper within a single equation.
To add climate variability to the system, the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
\begin{equation}\label{eq:nino3}
  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
\end{equation}
where $x(t)$ represents a climate time series which itself exhibits \gls{lfv} but not secular change.
This parameterization is analagous to the ``climate-informed'' approach described in several studies for estimating climate risk \citep{Delgado:2014ey,Merz:2014gf,Farnham:2018gs}.
Following \cref{eq:nino3}, when $\beta \neq 0$ there will be \gls{lfv}, and when $\gamma \neq 0$ there will be secular change.
The values of all parameters used for sampling climate risk are listed in the online supporting information for each of the three scenarios considered.

We represent the climate state variable $x(t)$ through an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} as well as a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
We model \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
To create an annual time series, we average the October-December values of the NINO3 index for each year.
Supplemental figure S1 shows a wavelet spectrum and time series plot of the resulting annual time series.
In the online supplemental information we consider an alternative parameterization of $\mu(t)$, which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.

@@

Once a synthetic streamflow sequence has been generated, we evaluate the identifiability and predictability of the dominant climate modes by fitting the sequence to statistical models and creating probabilistic projections of the future.
We use three well-studied statistical methods for future flood risk, each of which parameterizes time in a different way.
One is purely stationary, another captures \gls{lfv}, and the third captures secular change.
We choose these models for their interpretability and simplicity, rather than because of a belief that they are generally valid.
For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations.
Once a statistical model is fit to these observations, then $K=\num{1000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.

In the first case we fit a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
We refer to this model as ``LN2 Stationary.''
The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors \citep{gelman:2017,simpson:2017}.
The full model, including priors, is given in supplemental equation (S1).

Next, we modify this stationary model to incorporate secular change.
Many studies have done this by regressing certain parameters of the model on time \citep[see][for a comprehensive review]{Salas:2018ge}.
We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in supplemental equation (S2).
We refer to this model as ``LN2 Linear Trend.''
This model gives a lower bound on total informational uncertainty because it correctly represents the trend's known form, whereas in real-world analyses the form of the trend is unknown.

Finally, we explicitly model \gls{lfv} using a \gls{hmm}.
An \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (i.e.\@\xspace hidden) states $S(t)$ \citep{Rabiner:1986jk}.
The (unobserved) states evolve following a first-order Markov process, and the observed variable (e.g.\@\xspace streamflow) depends only on the underlying state.
\Glspl{hmm} have been widely used for modeling streamflow sequences \citep{Bracken:2016ba} and \gls{enso} \citep{rojohernandez:2017}.
We fit  streamflow sequences using a \gls{hmm} with two states.
The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution that is conditional only on the unobserved state variables.
This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
For simplicity, we fit only a two-state \gls{hmm} to each sequence.
Future floods are then estimated by simulating future states from the estimated transition matrix and then drawing $Q(t)$ conditional on the simulated state.

@@

Both estimation bias and estimation uncertainty affect the utility of a climate risk projection.
An instrument whose design was based on projections with overestimated variance or positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the public.

We evaluate both the estimation bias and estimation uncertainty.
For a given choice of $M$, $N$, and generating model, we compare the synthetic streamflow sequence's $N$-year ``historical record'' and the $K=1000$ posterior simulations of future flows.
The quantity $\hat{p}_T$, the estimated expected number of floods per year, is taken by calculating, for each of the $K$ posterior simulations, the number of exceedances of the flood design threshold, then dividing by $M$ to get exceedances per year.
We then compute the variance of these $K$ estimates.
We further calculate the bias of $\hat{p}_T$ by averaging it across the $K$ samples and comparing this to the number of times the $M$-year ``future period'' of the synthetic streamflow sequence exceeds the flood design threshold.
Since the ``observed'' number of flood exceedances from the generating model is inherently noisy for an $M$-year period, we average the bias and variance across $J=1000$ different streamflow sequences to compute expected values of both.
These sequences are generated with the same underlying parameters, but the specific synthetic NINO3 sequence (or set of Markov states) may differ between the $J$ sequences.

@@

\Cref{fig:flowchart} describes the experimental design.
We assess estimation bias and variance for three scenarios of future climate.
First, we consider an idealized scenario where only secular change is present in the system and \gls{lfv} is fully damped (``secular change only'').
Next, we consider the ``pre-industrial'' case where there is no secular change but \gls{lfv} modulates climate risk in time (``\acrlong{lfv} only'').
Finally, we consider a more realistic (though still idealized) case with both \gls{lfv} and secular change (``\acrlong{lfv} plus secular change'').
Model parameters for each scenario are given in the supplemental methods section of the online supporting materials.
